---
title: "Frequency"
author: "Gurudev Ilangovan"
date: "June 27, 2017"
output:
  html_document:
    toc: true
---

```{r echo='FALSE', include=FALSE, tidy=TRUE}
library(tidyverse)
library(ggthemr)
library(ggvis)
library(forcats)
library(RColorBrewer)

setwd("C:/Users/gurud/Google Drive/Research Assistantship/seq/Gurudev Ilangovan/Record Linkage/userstudy/pipeline/frequencies/")

fname_freq <- read_csv("./fname_freq.csv")
lname_freq <- read_csv("./lname_freq.csv")

starred_data_all <- read_csv("../data_output/all_starred_race.csv")

fname_starred <- 
  starred_data_all %>%
    select(`First Name`, n = FF)

lname_starred <- 
  starred_data_all %>%
  select(`Last Name`, n = LF)

custom <- define_palette(
  # swatch = c("black","blue","red","darkgreen","darkgrey","darkorange","pink"),
  swatch = c("#000000", brewer.pal(5,"Set1")),
  gradient = c(lower = "lightblue", upper = "darkblue")
)

ggthemr(custom)

```


```{r include=FALSE}
calculate_freq_thresh <- function(name_type,percentile){
   if(name_type == "fname"){
    freq <- fname_freq$n
    freq = freq[freq>5]
    return(quantile(freq,percentile))
  } else {
    lreq <- lname_freq$n
    lreq = lreq[lreq>5]
    return(quantile(lreq,percentile)  )
  }
}

calculate_rarity <- function(frequency,name_type = "fname", percentile = 0.5) {
  
  freq_threshold <- calculate_freq_thresh(name_type,percentile)  
  print(paste0("The threshold is ",freq_threshold))
    
  frequency <- tibble(freq = frequency)
  frequency %>% 
    mutate(rarity = as_factor(case_when(
      freq == 1 ~ "Unique",
      between(freq, 2, 5) ~ "Rare",
      between(freq, 5, freq_threshold) ~ "Common",
      freq>freq_threshold ~ "Frequent")),
      rarity = fct_relevel(rarity, "Unique", "Rare", "Common", "Frequent")) %>%
        pull(rarity) %>%
          return()
}

fname_freq <- 
  fname_freq %>%
    mutate(Rarity = calculate_rarity(n), name_type = "fname")

lname_freq <- 
  lname_freq %>%
    mutate(Rarity = calculate_rarity(n), name_type = "lname")


fname_starred <- 
  fname_starred %>%
    mutate(Rarity = calculate_rarity(n), name_type = "fname")

lname_starred <- 
  lname_starred %>%
    mutate(Rarity = calculate_rarity(n), name_type = "lname")

f_threshold <- calculate_freq_thresh("fname",0.5)
l_threshold <- calculate_freq_thresh("lname",0.5)
```
Current classification:

1. Unique  : 1
2. Rare    : 2 - 5
3. Common  : 6 - `r f_threshold`(for first name) / `r l_threshold`(for last name)
4. Frequent: Greater than `r f_threshold`(for first name) / `r l_threshold`(for last name)

The thresholds have been arrived at, using the following procedure:

- In the lookup table for the corresponding name, subset the frequencies greater than 5.
- With the remaining frequencies, calculate the frequency at the 50th percentile

##The original frequencies look like this:

###First Names:
```{r echo = FALSE}
fname_freq %>%
  mutate(src = ifelse(src == "B", "1703", "1304")) %>% 
  ggplot(aes(n)) + 
  facet_grid(~src) +
  geom_histogram(binwidth = 10, aes(fill = Rarity), alpha = 0.75)
```

###Last Names:
```{r echo = FALSE}
lname_freq %>%
  mutate(src = ifelse(src == "B", "1703", "1304")) %>% 
  ggplot(aes(n)) + 
  facet_grid(~src) +
  geom_histogram(binwidth = 10, aes(fill = Rarity), alpha = 0.8)
```

##The frequencies in the crafted data look like this:

###First Names:
```{r echo = FALSE}
fname_starred %>%
  ggplot(aes(n)) + geom_histogram(binwidth = 10,  alpha = 0.8, aes(fill = Rarity))
```

###Last Names:
```{r echo = FALSE}
lname_starred %>%
  ggplot(aes(n)) + geom_histogram(binwidth = 10,  alpha = 0.8, aes(fill = Rarity))
```

##Percentage of Frequencies in the lookup table vs crafted data
```{r echo=FALSE}
ggplot(fname_freq, aes(Rarity, fill = Rarity)) +
  geom_bar(aes(y = 100 * (..count..)/sum(..count..)), alpha = 0.7) +
  ylab("Percentage") + ggtitle("Lookup First Name")

ggplot(fname_starred, aes(Rarity, fill = Rarity)) +
  geom_bar(aes(y = 100 * (..count..)/sum(..count..)), alpha = 0.7) +
  ylab("Percentage") + ggtitle("Starred First Name")

ggplot(lname_freq, aes(Rarity, fill = Rarity)) +
  geom_bar(aes(y = 100 * (..count..)/sum(..count..)), alpha = 0.7) +
  ylab("Percentage") + ggtitle("Lookup Last Name")

ggplot(lname_starred, aes(Rarity, fill = Rarity)) +
  geom_bar(aes(y = 100 * (..count..)/sum(..count..)), alpha = 0.7) +
  ylab("Percentage") + ggtitle("Starred Last Name")


```


```{r}
fname_starred %>%
  count(n) %>%
    arrange((n)) %>%
      print(n = 100)


lname_starred %>%
  count(n) %>%
  arrange(desc(nn)) %>%
    print(n = 100)
```


```{r include=FALSE,eval=FALSE}

fname_starred %>%
  ggvis(~FF) %>% 
          layer_histograms(width = input_slider(min = 1,
                                                max = 100,
                                                value = 50))

```

